---
title: "Evaluator Workflow"
author: "David F. Severski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Evaluator Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Background

The first iterations of Evaluator were created as a part of a major healthcare 
organization's decision to shift its already mature risk assessment program 
from reliance on qualitative labels to a quantitative model that would support 
more precise comparison of competing projects. This organization was able to 
use statistical sampling to gain greater insight into its information risks, 
to meet HIPAA compliance obligations and to provide manager to board 
level business leaders with the data needed to drive decision making.

Since its creation, versions of Evaluator have been deployed both inside and 
outside the healthcare field.

# How to Use

The Evaluator toolkit consists of a series of processes implemented in 
the [R language](https://www.r-project.org/). Starting from an Excel workbook, 
risk data is imported and run through a simulation model to estimate 
the expected losses for each scenario. The results of these simulations are 
used to create a detailed analysis and a formal risk report. A starter 
analysis report, overview dashboard and sample 
[Shiny](https://shiny.rstudio.com/) application are all included in the toolkit.

Evaluator takes a domain-driven and framework-independent approach to strategic 
security risk analysis. It is compatible with ISO, COBIT, HITRUST CSF, 
PCI-DSS or any other model used for organizing an information security program. 
If you are able to describe the domains of your program and the controls and 
threat scenarios applicable to each domain, you will be able to use Evaluator!

## Instructions

This README does not define terms commonly used in an OpenFAIR analysis. 
While not a prerequisite, a review of OpenFAIR methodology and terminology is 
highly recommended. Familiarity with the R language is also very helpful.

Follow these seven steps to run the toolkit:

1. Define your controls and risk scenarios
2. Import the scenarios
3. Validate the data is ready for simulation
4. Encode the qualitative labels into quantiative parameters
5. Run the simulations
6. Summarize the simulation outputs
7. Analyze the results

Don't be intimidated by the process. Evaluator is with you at every step!

### Prepare the Environment

A working [R interpreter](https://www.r-project.org/) is required. 
Evaluator should work on any current version of R (v3.4.1 as of this writing) 
and on any supported platform (Windows, MacOS, or Linux). This vignette assumes 
the use of [RStudio IDE](https://www.rstudio.com/) for some presentation details, 
but the core Evaluator functions may be used with any IDE.

Obtain the Evaluator toolkit either from CRAN via `install.packages("evaluator")` or 
the current development version via `devtools::install_github('davidski/evaluator')`.

If you would like to work with the supplied sample data files, execute 
the following code:
```{r prepare_sample_environement, eval=FALSE}
library(dplyr)     # piping
library(readr)     # better CSV handling
library(evaluator) # core evaluator toolkit

# create default directories
input_directory <- "~/evaluator/data"
results_directory <- "~/evaluator/results"
if (!dir.exists(input_directory)) dir.create(input_directory, recursive = TRUE)
if (!dir.exists(results_directory)) dir.create(results_directory, recursive = TRUE)

# copy sample files
file.copy(system.file("extdata", "domains.csv", package = "evaluator"), 
          input_directory)
file.copy(system.file("extdata", "qualitative_mappings.csv", package="evaluator"), 
          input_directory)
file.copy(system.file("extdata", "risk_tolerances.csv", package = "evaluator"), 
          input_directory)
file.copy(system.file("survey", "survey.xlsx", package = "evaluator"),
          input_directory)
```

### Define Your Controls and Security Domains

Evaluator needs to know the domains of your security program. These are the 
major buckets into which you subdivide your program, typically including areas 
such as Physical Security, Strategy, Policy, Business Continuity/Disaster 
Recovery, Technical Security, etc. Out of the box, Evaluator comes 
with a demonstration model based upon the 
[HITRUST CSF](https://hitrustalliance.net/hitrust-csf/). If you have a 
different domain structure (e.g. ISO2700x, NIST CSF, or COBIT), you will
need to edit the `data/domains.csv` file to include the domain names and the 
domain IDs, and a shorthand abbreviation for the domain (such as POL for the 
Policy domain).

Indentifying the controls (or capabilities) and risk scenarios associated  with 
each of your domains is critical to the final analysis. The elements are 
documented in an Excel workbook. The workbook includes one tab per domain, with 
each tab named after the domain IDs defined in the previous step. Each tab 
consists of a controls table and a threats table.

#### Controls Table

The key objectives of each domain are defined in the domain controls table. 
While the specific controls will be unique to each organization, the 
sample spreadsheet included in Evaluator may be used as a model. It is best to
avoid copying every technical control from, for example, ISO 27001 or COBIT, 
since most control frameworks are too fine-grained to provide the high level 
overview that Evaluator delivers. In practice, 50 controls or less will 
usually be sufficient to describe organizations of up to one to two billion 
USD in size. Each control must have a unique ID and should be assigned a 
difficulty (DIFF) score that ranks the maturity of the control on a CMM 
scale from Initial (lowest score) to Optimized (best of class).

#### Threats Table

The threats table consists of the potential loss scenarios addressed by each 
domain of your security program. Each scenario is made up of a descriptive 
field that describes who did what to whom, the threat community that 
executed the attack (e.g. external hacktivist, internal workforce member, 
third party vendor), how often the threat actor acts upon your assets (TEF), 
the strength with which they act against your assets (TCap), the losses 
incurred (LM) and a comma-separated list of the control IDs that prevent 
the scenario.

### Import the Scenarios

To extract the spreadsheet into data files for further analysis, run 
`import_spreadsheet("PATH", domains)`. Evaluator will extract the data and 
save two comma seperated files in a `data` directory with the results. 

```{r import, eval=FALSE}
domains <-  readr::read_csv(file.path(input_directory, "domains.csv"))
system.file("survey", "survey.xlsx", package = "evaluator") %>% 
  import_spreadsheet(., domains, output_dir = input_directory)
```

### Validate the Data
Run the command `validate_scenarios()`. If there are data validation errors, 
the process will abort and an error message will be displayed. Correct the 
errors displayed, reimport, and repeat the validation process until there are 
no errors reported. 
 
```{r validate, eval=FALSE}
mappings <-  readr::read_csv(file.path(input_directory, "qualitative_mappings.csv"))
qualitative_scenarios <- readr::read_csv(file.path(input_directory, "qualitative_scenarios.csv"))
capabilities <- readr::read_csv(file.path(input_directory, "capabilities.csv"))
validate_scenarios(qualitative_scenarios, capabilities, domains, mappings)
```


### Encode the Data

```{r encode, eval = FALSE}
quantitative_scenarios <- encode_scenarios(qualitative_scenarios, 
                                           capabilities, mappings)
```

### Run the Simulations

Once the data is translated into quantitative scenarios and ready for 
simulation, run `run_simulations(quantitative_scenarios)`. By default, 
Evaluator puts each scenario through 10,000 individual simulated years, 
modelling how often the threat actor will come into contact with your assets,
the strength of the threat actor, the strength of your controls, and the 
losses involved in any situation where the threat strength exceeds your 
control strength. This simulation process can be computationally intense. The 
sample data set takes approximately 24 minutes to run on a 2016 MacBook Pro.

```{r simulate, eval = FALSE}
simulation_results <- run_simulations(quantitative_scenarios, 
                                      simulation_count = 100L)
save(simulation_results, file = file.path(results_directory, "simulation_results.rda"))
```

### Analyze the Results

Several analysis functions are provided, including a template for a technical 
risk report. Assuming that the input files have been placed in a `data` 
directory and the simulation results and summary files in a `results` directory, 
the risk report can be generated via 
`generate_report(input_directory, results_directory)`.
This will create a pre-populated risk report that identifies key scenarios 
and generates initial plots for to be used in creating a final analysis report. 
Other included report tools include `risk_dashboard(input_directory, results_directory)`.

For interactive exploration, run 
`explore_scenarios(input_directory, results_directory)` 
to launch a local copy of the Scenario Explorer application. The Scenario 
Explorer app may be used to view information about the individual 
scenarios and provides a sample overview of the entire program. 

For more in depth analysis, the following data files may be used directly 
from R or from external business intelligence programs such as Tableau:

```{r data_files, eval = FALSE, echo=FALSE}
tibble::tribble(
    ~"Data File", ~Purpose,
    "simulation_results.rda", "Full details of each simulated scenario",
    "scenario_summary.rda", "Simulation results, summarized at the scenario level",
    "domain_summary.rda", "Simulation results, summarized at the domain level"
) %>% pander::pander(., justify = "left")
```

```{r analyze, eval=FALSE}
# summarize
scenario_summary <- summarize_scenarios(simulation_results)
domain_summary <- summarize_domains(simulation_results, domains)

# or to save the summary files directly to disk
summarize_to_disk(simulation_results = simulation_results, domains = domains, results_directory)

# define risk tolerances
risk_tolerances <- system.file("extdata", "risk_tolerances.csv", 
                               package="evaluator") %>% read_csv()

# Explorer
explore_scenarios(input_directory, results_directory)

# Sample Report
generate_report(input_directory, results_directory, "report.pdf", format = "pdf") %>% 
  rstudioapi::viewer()

# Risk Dashboard
risk_dashboard(input_directory, results_directory) %>% rstudioapi::viewer()
```

Or, to view that same report as a Word document for editing, use 
`generate_report(input_directory, results_directory, format = "word")`.

If you would rather work on the source RMarkdown file or direct editing and 
graphics tweaking, use `system.file("rmd, "analyze_risk.Rmd", package = "evaluator")` 
to find the location of the native template on your system.

## Advanced Customization

Evaluator makes several assumptions to get you up and running as quickly as 
possible. Advanced users may implement several different customizations 
including:

- Risk tolerances - Organizational risk tolerances at a "medium" and "high" 
level are defined in `risk_tolerances.csv`. Risk tolerances are the
aggregate economic loss thresholds defined by your organization. These are not 
necessarily the same as the size of potential losses from individual 
scenarios. A good proxy for risk tolerance is the budget authority implemented 
in your organization. The size of purchase signoff required at the executive 
level is generally a good indicator of the minimum floor for high risk 
tolerance.
- Qualitative mappings - The translation of qualitative labels such as 
"Frequent" threat events and "Optimized" controls are defined in 
`qualitative_mappings.csv`. The values in this mapping may be changed 
but they must use lowercase and agree with the values used in the survey 
spreadsheet. Changing the number of levels used for any qualitative label 
(e.g. changing High/Medium/Low to High/Medium/Low/VeryLow) is unsupported.
- Styling - Look and feel (fonts, colors, etc.) is defined in the 
`styles/html-styles.css` and `styles/word-styles-reference.docx` files.

# Where to Go From Here

While Evaluator is a powerful tool, it does not explicitly attempt to address 
complex analysis of security risks, interaction between risk scenarios, 
rolling up multiple levels of risk into aggregations, modelling secondary 
losses or other advanced topics. As you become more comfortable with 
quantitative risk analysis, you may wish to dive deeper into these areas 
(and I hope you do!).

Commercial Software

- [RiskLens](http://www.risklens.com/), founded by the original creator of the 
FAIR methodology

Blogs/Books/Training

- Russell C. Thomas's excellent and provocative blog post on systemic [Risk Management](http://exploringpossibilityspace.blogspot.com/2013/08/risk-management-out-with-old-in-with-new.html)
- [Measuring and Managing Information Risk](https://smile.amazon.com/gp/product/0124202314)
- [OpenFAIR certification](http://www.opengroup.org/certifications/openfair)
- [Hubbard Decision Research calibration training](https://www.hubbardresearch.com/training/)

Associations

- [FAIR Institute](http://www.fairinstitute.org/)
- [Society of Information Risk Analysts (SIRA)](https://www.societyinforisk.org/)
